# Description of the data

### Introduction

In this section, we'll take a closer look at various details about our heart disease dataset and provide a detailed explanation of the dataset.

### Loading the data

```{r}
#| message: false
library(tidyverse)
```

```{r}
# Loading the clean data
data <- read_tsv("../data/02_dat_clean.tsv")
data
```

### Dataset Overview

After loading the dataset into R, it revealed 14 variables and 303 observations. It provides key information, including details about age, gender, and various aspects of heart health of certain patients. The 'goal' field indicates the presence of heart disease, ranging from 0 (no presence) to 4. Also, patient names and social security numbers were replaced with dummy values for privacy. In the following, we will take a deeper look at the variable information to make the resulting interpretation of the dataset easier.

### Variable Information

The upcoming variable documentation consists of two different sub-parts. The documentation will be split into 'categorical variables' and 'continuous variables.' 'Categorical variables' contain a finite number of categories or distinct groups; for example, categorical predictors include gender, material type, and payment method. 'Continuous variables' are numeric variables with an infinite number of values between any two points, such as the length of a part or the date and time a payment is received.

#### Categorical variables

We have 7 categorical variables in the data:

+----------------------------------+----------------------------------------------------------------------------------+
| Variable Name                    | Description                                                                      |
+==================================+==================================================================================+
| sex                              | sex (male or female)                                                             |
+----------------------------------+----------------------------------------------------------------------------------+
| cp                               | chest pain type                                                                  |
|                                  |                                                                                  |
|                                  |  1. typical angina                                                               |
|                                  |                                                                                  |
|                                  | 2. atypical angina                                                               |
|                                  |                                                                                  |
|                                  | 3. non-anginal pain                                                              |
|                                  |                                                                                  |
|                                  | 4. asymptomatic                                                                  |
+----------------------------------+----------------------------------------------------------------------------------+
| fbs                              | fasting blood sugar \> 120 mg/dl                                                 |
+----------------------------------+----------------------------------------------------------------------------------+
| restecg                          | resting electrocardiographic results                                             |
|                                  |                                                                                  |
|                                  |  1. normal                                                                       |
|                                  |                                                                                  |
|                                  | 2. having ST-T wave abnormality                                                  |
|                                  |                                                                                  |
|                                  | 3. showing probable or definite left ventricular hypertrophy by Estes' criteria  |
+----------------------------------+----------------------------------------------------------------------------------+
| exang                            |                                                                                  |
+----------------------------------+----------------------------------------------------------------------------------+
| slope                            |                                                                                  |
+----------------------------------+----------------------------------------------------------------------------------+
| thal                             |                                                                                  |
+----------------------------------+----------------------------------------------------------------------------------+

### Continuous variables

| Variable Name | Description |
|---------------|-------------|
| age           |             |
| trestbps      |             |
| chol          |             |
| thalach       |             |
| oldpeak       |             |
| ca            |             |
| num           |             |

### Categorical variables

There are 9 categorial variables in the data:

-   sex

-   cp

-   fbs

-   restecg

-   exang

-   slope

-   ca

-   thal

-   num

The cp variable describes the chest pain of each observation. There are four categories in total:

-   Typical angina

-   Atypical angina

-   Non-anginal pain

-   Asymptomatic

Typical angina, also referred to as stable angina, is a pain which lasts a couple of minutes. It occurs in a patterns, and is induced by fx. stress or exercise.

Atypical angina, also referred to as unstable angina, is different from typical angina by lasting longer and not occuring in patterns.

@https://www.nhlbi.nih.gov/health/angina/types

The following table shows the number of observations from each category.

```{r}
data |>
  group_by(cp) |>
  count()
```

The num variable is a scale of numbers from 0 to 4. 0 means that the patient has not been diagnosed with a heart disease, while 1-4 indicates the type of heart disease.

### Continuous variables

The data contains five continuous variables:

-   age

-   trestbps

-   chol

-   thalach

-   oldpeak

The trestbps variable contains information on the resting blood pressure of each observation in mm Hg, on admission to the hospital. It ranges from 94 to 200 mm Hg, and has a mean of 131.7.

The chol variable contains information on the serum cholestoral of each observation in mg/dl. It ranges from 126 to 564, and has a mean of 246.7.

The thalach varibale contains information on the maximum heart rate achieved by each observation. It ranges from 71 to 202, and has a mean of 149.6.

```{r}
data |>
  select(trestbps, chol, thalach) |>
  summary()
```
